Stochastic Approximation Monte Carlo with a Dynamic Update Factor (SAD)

### 2019-03-07 papers/histogram/sad.tex
edit to two paragraphs of sad paper

\section{Introduction}
Over the past several decades a number of flat histogram Monte-Carlo
simulation algorithms have been developed which calculate the
thermodynamic properties of various systems for all temperatures.  The
development began with the original histogram method, which used a
single canonical Monte Carlo simulation to predict properties for
nearby temperatures~\cite{ferrenberg1988new}.  For large simulations
this approach is limited to a narrow temperature range because a single
canonical simulation explores only a small range of energies.  This led
to a variety of ``broad'' (or ``flat'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
wang2001determining, wang2001efficient, trebst2004optimizing}, which
attempt to explore a wide range of energies.  These methods also
benefit in that they are unlikely to be trapped in a low enetropy state.

Wang and Landau introduced an algorithm that uses an update factor and
a statistical histogram to compute the density of states of a given
system~\cite{wang2001determining, wang2001efficient}.  While the method
is incredibly powerful, it has several user-defined
parameters~\cite{landau2004new}.  This can make its application to a
variety of systems something of an art-form since the user needs to
determine the ideal parameters for the particular system being studied.
Also, detailed balance is violated (although only for short periods of
time), ensuring that convergence is not guaranteed.  This raises
several questions such as when the method does converge, at what rate will
the method converge to the correct density of states. Also, can the
user appropriately decide what parameters to choose so that the
algorithm solves the problem in the most ideal way.

>>>>

\section{Introduction}
Over the past several decades a number of flat histogram Monte-Carlo
simulation algorithms have been developed which calculate the
thermodynamic properties of various systems for all temperatures.  The
development began with the original histogram method, which used a
single canonical Monte Carlo simulation to predict properties for
nearby temperatures~\cite{ferrenberg1988new}.  For large systems this
approach is limited to a narrow temperature range because a single
canonical simulation explores only a small range of energies.  This
led to a variety of ``broad'' (or ``flat'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
  wang2001determining, wang2001efficient, trebst2004optimizing}, which
attempt to explore a wide range of energies.  Another benefit, in
contrast with low-temperature canonical Monte Carlo, is that these
approaches cannot be trapped in a local energy minimum.

Wang and Landau introduced an algorithm (WL) that uses an update
factor and a statistical histogram to compute the density of states of
a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has a few disadvantage.
Firstly, it requires the user to select a number of parameters,
including the range of energies to be studied~\cite{landau2004new}.
This adds an additional hurdle to its application to a variety of
systems.  The simulation violates detailed balance, with the size of
the violation decreasing with time, which complicates convergence
analysis.  In fact, the error in a WL computation the error has been
demonstrated to saturate at a non-zero value, i.e. the method does not
converge to the true density of states~\davidsays{Find as many
  citations as we can! Presumably some from the following paragraph?}.
